### 2021-10-21 paper/manuscript.Rmd
Results and script for reproduction. Images are in seperate .rtf file, but not pushed due to size (67 MB).

#introduction
**intro to meta-analysis**
Meta-analysis is a statistical method which utilizes several tools to synthesize the data of multiple studies on the same topic, with the purpose of finding a result that is more trustworthy. What meta-analysis does is simply weighting all the observed effect sizes of the individual studies and averaging them to one summary effect. Although this explanation is a bit too simplistic, in essence this is what meta-analysis is about. Meta-analysis assigns weights to each individual study based on different assumptions which are set in advance. These weights determine to what extent an individual study takes part in the eventual summary effect.

@@

#Results
There were 20 cases out of 388800 that had missing values on all metrics for the RMA algorithm. Closer inspection showed that both the cubic and exponential model each contributed 10 times to the missing values and only when 2, 3 or 6 moderators were taken up in the model. However, since this makes up 0.005% of the data, the missing values were chosen to be omitted from further analysis.

*Predictive performance*
Because the densities for the $R^{2}$ values were either bimodal or skewed, it was chosen to use the median $R^{2}$ as the metric for predictive performance, rather than the mean. The spread of the data was described using the Mean Absolute Deviation [MAD], rather than the standard deviation. It was found that the Horseshoe, Lasso and RMA algorithm performed similarly overall, $Med_{hs}$ = 0.51, MAD = 0.36; $Med_{Lasso}$ = 0.50, MAD = 0.37; $Med_{RMA}$ = 0.50, MAD = 0.37. The random forest algorithm performed worst $Med_{rf}$ = 0.35, MAD = 0.38.

To determine the effect of the design factors on predictive performance of all algorithms, four separate ANOVA’s were performed. The effect size $η^{2}$ per condition per algorithm, including interactions can be found in table 1.

Not too surprisingly, It was found that the true effect size $β$ had the largest effect on the predictive $R^{2}$ for all algorithms. As $\beta$ increased, the performance of all algorithms increased as well. However, $\beta$ did seem to interact with the model that was estimated, being either a linear, two-way interaction, cubic or an exponential model. The image of the interaction is shown in image 1. For the exponential and two-way interaction model, the increase of $R^{2}$ slows for every higher value for $\beta$. The cubic model only slows the increase when the es goes from 0.5 to 0.8, while the steepest increase for the linear model is when $\beta$ increases from 0.2 to 0.4. Also noteworthy is that the two-way interaction model seems to stagnate as the effect size goes up to 0.4 and onwards, while the other models do keep increasing. There seemed to be little difference in performance between Horsehoe, Lasso and RMA, while MetaForest performed worst.

The second largest marginal effect was that of the estimated model. It seemed that all algorithms performed best under the two-way interaction model, followed by a similar performance on the cubic and exponential models. All algorithms seemed to perform worst for the linear model. There again is little difference in performance between the Pema algorithms and RMA, altough MetaForest performs worst. Image 2 shows the relationship.

There also was a moderate interaction effect between the estimated model and the amount of skewness of the data [$\alpha$], especially for the Pema algorithms. Again, Pema and RMA algorithms performed best, followed by MetaForest in all conditions. Most obvious to note is that all, except the linear model seem to benefit when the $R^{2}$ was estimated on more skewed data, although the algorithms do perform worse when the model is cubic and $\alpha$ goes from 5 to 10. The two-way interaction model performs best overall, but is getting caught up by the exponential and cubic model as $\alpha$ increases. Image 3 shows the relationship.

@@

The mean sample size per study [$\bar{n}$] also had a moderate effect. For all algorithms the effect of $\bar{n}$ was positively linearly related to $R^{2}$ with the Pema and RMA algorithms performing best and Metaforest performing worse. Image 5 shows the relationship.

@@

*Estimating residual heterogeneity* ($τ^{2}$)
The ability of the algorithms to correctly estimate $τ^{2}$ was operationalized by subtracting the true value for $\tau^{2}$ from the $\tau^{2}$ estimated by the algorithms. Again, the median and MAD were used as metrics for performance. The RMA algorithm showed the best results, $Med_{rma}$ = 0.02, MAD = 0.06, followed by the Metaforest algorithm $Med_{mf}$ = 0.09, MAD = 0.13. The Pema algorithms performed worst $Med_{lasso}$ = 0.23, MAD = 0.18; $Med_{hs}$ = 0.23, MAD = 0.17. The finding that all medians are positive implies that all algorithms have a bigger tendency to overestimate $\tau^{2}$ than underestimate it. One general comment to make is that uncertainty of the estimates generally increased as $\Delta\tau^{2}$ also increased. This implies that there is more variation in performance as median performance worsens.

To determine the effect of the design factors on estimated $\tau^{2}$ of all algorithms, four separate ANOVA’s were performed. The effect size $\eta^{2}$ per condition per algorithm, including interactions can be found in table 2.

The estimated model had one of the biggest effect on the correct estimation of $τ^{2}$. This is mainly because the algorithms overestimate $\tau^{2}$ quite severely when the model contains two-way interactions. Image 8 shows the marginal relationship of the estimated model on $\Delta\tau^{2}$ . It becomes more clear why this overestimation occurs when showing the interaction between the effect size and the model estimated on $\Delta\tau^{2}$ shown in image 9. First note the general trend which seems to be that for all models, except the linear, $\tau^{2}$ gets more overestimated the higher the true effect size is. This is not the case for the linear algorithm, where the effect of $\beta$ on $\Delta\tau^{2}$  seems close to zero, except for MetaForest. However, note the scales for the y-axes. While the cubic, linear and stay well within a confined interval, the algorithms severely overestimate $\tau^{2}$ when the model contains two-way interaction terms. Especially Metaforests overestimation is substantial when $\beta$ = 0.8 and the estimated model contains two-way interactions with $Med_{hs}$ = 2.92, $MAD$ = 2.37, although all other algorithms also have a $\Delta\tau^{2} > 1$ in these conditions. Interestingly, the Pema algorithms even outperform the RMA algorithm within these conditions. The marginal effects of $\beta$ on $\Delta\tau^{2}$  are shown in image 10. MetaForest seems most affected by the increase in $\beta$, but in general performs better than the Pema algorithms when $\beta$ < 0.8. The RMA algorithm performs best overall.

The marginal effect of the skewness of the data $\alpha$ on $\Delta\tau^{2}$  seems rather minimal. There is a slight decrease in $\Delta\tau^{2}$  as $\alpha$ increases, but it is hardly noticeable. However, this decrease becomes more clear when we add the interaction with the estimated model. Image 12 shows the conditional relationship. The algorithms seem rather unaffected by $\alpha$ for the linear model, and a small general decrease in $\Delta\tau^{2}$  as $\alpha$ increases can be seen in the exponential model. When the cubic model is estimated however, the algorithms seem to benefit as $\alpha$ increases, while for the two-way interaction model, $\Delta\tau^{2}$  first increases as $\alpha$ increases, and afterwards decreases. RMA performs best, followed by Metaforest. The Pema algorithms perform similar, but worst.

@@

The effect of the number of moderators on $\Delta\tau^{2}$  was not that large either. A small increase in $\Delta\tau^{2}$ can be seen in the RMA and Metaforest algorithm as the number of moderators increase, which is not found for the Pema algorithms. However, A small note that should be made is that the Metaforest algorithm does substantially increase in $\Delta\tau^{2}$  as more moderators are added when the estimated model contains two-way interactions. Image 13 shows the interaction between the number of moderators and the estimated model on $\Delta\tau^{2}$ .

The number of studies used in the training data $\kappa$ only had a very substantial effect for the Metaforest algorithm. For Metaforest, the $\Delta\tau^{2}$  decreased quite rapidly if $\kappa$ increased, especially when the two-way interaction model was estimated. For the other algorithms, decreasing $\kappa$ seemed to have little to know effect on correctly estimating the residual heterogeneity. Image 14 shows the relationship.

@@

*Variable selection*
**currently working on**

>>>>

# introduction
**intro to meta-analysis**
Meta-analysis is a statistical method which utilizes several tools to synthesize the data of multiple studies on the same topic, with the purpose of finding a result that is more trustworthy. What meta-analysis does is simply weighting all the observed effect sizes of the individual studies and averaging them to one summary effect. Although this explanation is a bit too simplistic, in essence this is what meta-analysis is about. Meta-analysis assigns weights to each individual study based on different assumptions which are set in advance. These weights determine to what extent an individual study takes part in the eventual summary effect.

@@

# Results
There were 20 cases out of 388800 that had missing values on all metrics for the RMA algorithm. Closer inspection showed that both the cubic and exponential model each contributed 10 times to the missing values and only when 2, 3 or 6 moderators were taken up in the model. However, since this makes up 0.005% of the data, the missing values were chosen to be omitted from further analysis. Another thing is that the two-way interaction model only has results when there were either 3,4 or 7 moderators taken up, while the other models only have results when there were 2,3 or 6 moderators. The effects of the moderators is therefore difficult to compare between the two-way interaction and the other models.

*Predictive performance*
Because the densities for the $R^{2}$ and $MSE$ values were skewed, it was chosen to use the median as the metric for predictive performance, rather than the mean. The spread of the data was described using the Mean Absolute Deviation [MAD], rather than the standard deviation. It was found that the Horseshoe, Lasso and RMA algorithm performed similarly overall, $R^2_{hs} = 0.51 \pm 0.36$, $MSE_{hs} = 0.21 \pm 0.18$ ; $R^2_{Lasso} = 0.50 \pm 0.37$, $MSE_{Lasso} = 0.21\pm 0.19$; $R^2_{RMA} = 0.50 \pm 0.37$, $MSE_{RMA} = 0.22 \pm 0.23$ . The random forest algorithm performed worst on $R^2$ $R^2_{rf} = 0.35 \pm 0.38$, $MSE_{rf} = 0.22 \pm 0.19$

To determine the effect of the design factors on predictive performance $R^2$ of all algorithms, four separate ANOVA’s were performed. The effect size $η^{2}$ per condition per algorithm, including interactions can be found in table 1. Do note that the ANOVA's were performed with the normality assumption violated. The estimates serve as a guidance, rather than an non-contestable result.

Not too surprisingly, It was found that the true effect size $β$ had the largest effect on the predictive $R^{2}$ for all algorithms. As $\beta$ increased, the performance of all algorithms increased as well. However, $\beta$ did seem to interact with the model that was estimated, being either a linear, two-way interaction, cubic or an exponential model. The image of the interaction is shown in image 1. For the exponential and cubic model, the increase of $R^{2}$ slows for every higher value for $\beta$. The two-way interaction model only slows the increase when the es goes from 0.5 to 0.8, while the steepest increase for the linear model is when $\beta$ increases from 0.2 to 0.4. Also noteworthy is that the cubic model seems to stagnate as the effect size goes up to 0.4 and onwards, while the other models do keep increasing. There seemed to be little difference in performance between Horsehoe, Lasso and RMA, while MetaForest performed worst.

The second largest marginal effect was that of the estimated model. It seemed that all algorithms performed best under the cubic model, followed by a similar performance on the two-way interaction and exponential models. All algorithms seemed to perform worst for the linear model. There again is little difference in performance between the Pema algorithms and RMA, altough MetaForest performs worst. Image 2 shows the relationship.

There also was a moderate interaction effect between the estimated model and the amount of skewness of the data [$\alpha$], especially for the Pema algorithms. Again, Pema and RMA algorithms performed best, followed by MetaForest in all conditions. Most obvious to note is that all, except the linear model seem to benefit when the $R^{2}$ was estimated on more skewed data, although the algorithms do perform worse for the interaction model and $\alpha$ goes from 5 to 10. The cubic model performs best overall, but is getting caught up by the exponential and interactoin model as $\alpha$ increases. Image 3 shows the relationship.

@@

The mean sample size per study [$\bar{n}$] also had a moderate effect. For all algorithms the effect of $\bar{n}$ was positively linearly related to $R^{2}$ with the Pema and RMA algorithms performing best and Metaforest performing worst. Image 5 shows the relationship.

@@

*Estimating residual heterogeneity* ($τ^{2}$)
The ability of the algorithms to correctly estimate $τ^{2}$ was operationalized by subtracting the true value for $\tau^{2}$ from the $\tau^{2}$ estimated by the algorithms. Again, the median and MAD were used as metrics for performance. The RMA algorithm showed the best results, $\Delta\tau^2_{RMA} = 0.02 \pm 0.06$, followed by the Metaforest algorithm $\Delta\tau^2_{MF} = 0.09 \pm 0.13$. The Pema algorithms performed worst $\Delta\tau^2_{Lasso} = 0.23 \pm 0.18$; $\Delta\tau^2_{hs} = 0.23 \pm 0.17$. The finding that all medians are positive implies that all algorithms have a bigger tendency to overestimate $\tau^{2}$ than underestimate it. One comment to make is that uncertainty of the estimates generally increased as $\Delta\tau^{2}$ also increased. This implies that there is more variation in performance as median performance worsens.

To determine the effect of the design factors on estimated $\tau^{2}$ of all algorithms, four separate ANOVA’s were performed. The effect size $\eta^{2}$ per condition per algorithm, including interactions can be found in table 2. Again, the assumption of normality was violated.

The estimated model had one of the biggest effect on the correct estimation of $τ^{2}$. This is mainly because the algorithms overestimate $\tau^{2}$ quite severely when the model contains cubic relationships. Image 8 shows the marginal relationship of the estimated model on $\Delta\tau^{2}$ . It becomes more clear why this overestimation occurs when showing the interaction between the effect size and the model estimated on $\Delta\tau^{2}$ shown in image 9. First note the general trend which seems to be that for all models, except the linear, $\tau^{2}$ gets more overestimated the higher the true effect size is. This is not the case for the linear algorithm, where the effect of $\beta$ on $\Delta\tau^{2}$  seems close to zero, except for MetaForest. However, note the scales for the y-axes. While the interaction, linear and stay well within a confined interval, the algorithms severely overestimate $\tau^{2}$ when the model contains cubic terms. Especially Metaforests overestimation is substantial when $\beta$ = 0.8 and the estimated model is cubic with $\Delta\tau^{2}_{MF} = 2.92 \pm 2.37$, although all other algorithms also have a $\Delta\tau^{2} > 1$ in these conditions. Interestingly, the Pema algorithms even outperform the RMA algorithm within these conditions. The marginal effects of $\beta$ on $\Delta\tau^{2}$  are shown in image 10. MetaForest seems most affected by the increase in $\beta$, but in general performs better than the Pema algorithms when $\beta$ < 0.8. The RMA algorithm performs best overall.

The marginal effect of the skewness of the data $\alpha$ on $\Delta\tau^{2}$  seems rather minimal. There is a slight decrease in $\Delta\tau^{2}$  as $\alpha$ increases, but it is hardly noticeable. However, this decrease becomes more clear when we add the interaction with the estimated model. Image 12 shows the conditional relationship. The algorithms seem rather unaffected by $\alpha$ for the linear model, and a small general decrease in $\Delta\tau^{2}$  as $\alpha$ increases can be seen in the exponential model. When the interaction model is estimated however, the algorithms seem to benefit as $\alpha$ increases, while for the cubic model, $\Delta\tau^{2}$  first increases as $\alpha$ increases, and afterwards decreases. RMA performs best, followed by Metaforest. The Pema algorithms perform similar, but worst.

@@

The effect of the number of moderators on $\Delta\tau^{2}$  was not that large either. A small increase in $\Delta\tau^{2}$ can be seen in the RMA and Metaforest algorithm as the number of moderators increase, which is not found for the Pema algorithms. However, A small note that should be made is that the Metaforest algorithm does substantially increase in $\Delta\tau^{2}$  as more moderators are added when the estimated model is cubic. Image 13 shows the interaction between the number of moderators and the estimated model on $\Delta\tau^{2}$ .

The number of studies used in the training data $\kappa$ only had a substantial effect for the Metaforest algorithm. For Metaforest, the $\Delta\tau^{2}$  decreased quite rapidly if $\kappa$ increased, especially when the cubic model was estimated. For the other algorithms, decreasing $\kappa$ seemed to have little to no effect on correctly estimating the residual heterogeneity. Image 14 shows the relationship.

@@

*Variable selection*
To determine the extent to which the algorithms could perform correct variable selection, the proportion true positives [TP] and true negatives [TN] were calculated. There were no differences in variables selected by Highest Density Intervals or Confidence Intervals for both Lasso and Horshoe and so for both algorithm it did not matter whether the Highest Density Interval or Confidence Interval were analyzed. The mean proportions along with the standard deviations on TP and TN are provided. It was found that Metaforest had the highest proportion true positives: $TP_{mf} = 0.98 \pm 0.15$, closely followed by RMA: $TP_{RMA} = 0.96 \pm 0.21$. Horshoe performed slightly better than Lasso;  $TP_{hs} = 0.91 \pm 0.29$; $TP_{Lasso} = 0.89 \pm 0.31$. As for TN, it was found that the pema algorithms performed best: $TN_{hs} = 0.93 \pm 0.16$, $TN_{Lasso} = 0.93 \pm 0.17$, followed by RMA: $TN_{RMA} = 0.89 \pm 0.20$. MF performed worst by a large margin: $TN_{mf} = 0.50 \pm 0.35$ 
Perfomance on TP and TN were very high for all algorithms, with all mean proportions exceeding .89. One exception was the performance on TN for MetaForest, which mean was 0.50. This implies that MetaForest had issues excluding irrelevant moderators from the models. Plots were inspected to determine the effects of the design factors on the proportions. Something different to note is that there seem to be little marginal effects of the design factors on TN, while TP is more affected. It implies that the algorithms are more robust in excluding irrelevant moderators than in including relevant moderators.

Firstly, $\kappa$ only had a marginal positive effect on TP. As $\kappa$ increased, TP also increased. This implies that the ability for all algorithms to attribute positive relevance to relevant moderators increased as the number of studies in the training dataset increased. MetaForest had the highest TP, followed by RMA and lastly Horshoe and Lasso. The increase in TP for higher values of $\kappa$ was steeper for the pema algorithms, however. There also seemed to exist an interaction of $\kappa$ with the estimated model. For the linear and interaction model, the relationship of $\kappa$ seemed relatively linear. At $\kappa = 20$ the TP was relatively low for the algorithms, compared to the cubic model where TP starts at .98 and converges to 1 as $\kappa$ increased. This latter relationship was also found for the exponential model, although the TP at $\kappa = 20$ was slighlty lower. Image 16 shows the interaction.

The mean observations per study had a roughly positive linear relationship with TP for all algorithms. MetaForest performed best, followed by RMA, while Horseshoe and Lasso performed worst. Image 17 shows the relationship.

The true effect size $\beta$ had an interaction effect with the model estimated on TN. Only for the interaction model, TN decreased as $\beta$ increased, For the other models, TN remained stable. This could be because the interaction model was only fitted when there were 3,4 or 7 moderators, while for the other models, only 2,3 or 6 moderators were used. Image 18 shows the relationship. The effect of $\beta$ on TP was positive; as $\beta$ increased, TP increased too.

The true $\tau^2$ had a negative effect on TP, while skewness parameter $\alpha$ seemed to have little effect of both TP and TN. Image 19 shows the marginal relationship of $\tau^2$ on TP.

Finally, there seemed to be an effect of number of moderators on TN only for the interaction model. The TN increased as the number of moderators did. Image 20 shows the relationship. The relationship was reversed for TP and was found for all algorithms, i.e. TP decreased as the number of moderators increased.
